---
title: "Lab 05 - Wrangling spatial data"
author: "Alex Connolly"
date: "IFeb 21 2022"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r alaska}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)

```
3 locations in Alaska

### Exercise 2


```{r lqak}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```

2



### Exercise 3


```{r joining}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak

```

### Exercise 4

6 observations, the variables are Address, city, zip, longitude, and latitude of both x and y (seperately) and state



### Exercise 5

```{r haversine}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```


### Exercise 6

```{r distance}
dn_lq_ak <- dn_lq_ak %>%
  mutate(dn_lq_ak, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_ak
```
### Exercise 7

```{r mindist}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_ak_mindist
```

### Exercise 8
```{r plot}
dn_lq_ak%>%
  ggplot(mapping = aes(x=address.x, y=distance)) + geom_point()
```

```{r plot2}
dn_lq_ak_mindist %>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```
As you can see, each Dennys has a La Quinta fairly close to it. Zooming in, you can see how one Dennys is even closer to a La Quinta, the one on 2900 Denali. 

### Exercise 9

```{r NorthCarolina}
dn_nc <- dennys %>%
  filter(state == "NC")
nrow(dn_nc)

```
28 locations in North Carolina



```{r lqnc}
lq_nc <- laquinta %>%
  filter(state == "NC")
nrow(lq_nc)
```

12 La Quinta locations in NC


```{r joining2}
dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")
dn_lq_nc

```

336 observations, the variables are Address, city, zip, longitude, and latitude of both x and y (separately) and state



```{r distance2}
dn_lq_nc <- dn_lq_nc %>%
  mutate(dn_lq_nc, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_nc
```

```{r mindist2}
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_nc_mindist
```

```{r plot3}
dn_lq_nc%>%
  ggplot(mapping = aes(x=address.x, y=distance)) + geom_point()
```
As you can see here, there are some La Quintas that are close to Dennys, but some far away. 

```{r plot11}
dn_lq_nc_mindist%>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```
So, lets look at the closest for each one. It does seem as each of them have one that is fairly close, over half within 50. 

### Exercise 10


```{r Texas}
dn_tx <- dennys %>%
  filter(state == "TX")
nrow(dn_tx)

```
200 Dennys locations in Texas



```{r lqtx}
lq_tx <- laquinta %>%
  filter(state == "TX")
nrow(lq_tx)
```

237 La Quinta locations in TX


```{r joining3}
dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")
dn_lq_tx

```

47,400 observations, the variables are Address, city, zip, longitude, and latitude of both x and y (separately) and state



```{r distance3}
dn_lq_tx <- dn_lq_tx %>%
  mutate(dn_lq_tx, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_tx
```

```{r mindist3}
dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_tx_mindist
```

```{r plot10}
dn_lq_tx%>%
  ggplot(mapping = aes(x=address.x, y=distance)) + geom_boxplot()
```
With how many observations were here, the scatter plot we have been using was not that helpful. This box and whisker plot, while tough to read, shows the sheer number of obserations, and how some are close but a lot are very far away from some Laquitas. But this is not that helpful since there are so many of each. So let's look at closest. 



```{r plot4}
dn_lq_tx_mindist%>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```
As you can see here, it does appear as if there is a La Quinta very close to almost every Dennys, only a handful were farther away. 
### Exercise 11

```{r Pennsylvania}
dn_pa <- dennys %>%
  filter(state == "PA")
nrow(dn_pa)

```
40 locations in Pennsylvania



```{r lqpa}
lq_pa <- laquinta %>%
  filter(state == "PA")
nrow(lq_pa)
```

10 La Quinta locations in PA


```{r joining4}
dn_lq_pa <- full_join(dn_pa, lq_pa, by = "state")
dn_lq_pa

```

400 observations, the variables are Address, city, zip, longitude, and latitude of both x and y (separately) and state



```{r distance4}
dn_lq_pa <- dn_lq_pa %>%
  mutate(dn_lq_pa, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_pa
```

```{r mindist4}
dn_lq_pa_mindist <- dn_lq_pa %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_pa_mindist
```

```{r plot7}
dn_lq_pa%>%
  ggplot(mapping = aes(x=address.x, y=distance)) + geom_point()
```
There are a lot here. Some seem close, but not as close as the other states. Lets take a closer look

```{r plot8}
dn_lq_pa_mindist%>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```
Looking at the closest, there are some that are close. But this is probably the worst of the four states. 
### Exercise 12

I would say that Texas it probably was the most applicable. Pretty much every Dennys had a LaQuinta super close. The Alaska ones did too - very close - but there was only 3, so the joke is funnnier in Texas. NC and PA were not nearly as good. 
