---
title: "Lab 05 - Wrangling spatial data"
author: "Alex Connolly"
date: "IFeb 21 2022"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r alaska}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)

```
3 locations in Alaska

### Exercise 2


```{r lqak}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```

2



### Exercise 3


```{r joining}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak

```

### Exercise 4

6 observations, the variables are Address, city, zip, longitude, and latitude of both x and y (seperately) and state



### Exercise 5

```{r haversine}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```


### Exercise 6

```{r distance}
dn_lq_ak <- dn_lq_ak %>%
  mutate(dn_lq_ak, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_ak
```
### Exercise 7

```{r mindist}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_ak_mindist
```

### Exercise 8
```{r plot}
dn_lq_ak%>%
  ggplot(mapping = aes(x=address.x, y=distance)) + geom_point()
```

```{r plot2}
dn_lq_ak_mindist %>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```
As you can see, each Dennys has a La Quinta fairly close to it. Zooming in, you can see how one Dennys is even closer to a La Quinta, the one on 2900 Denali. 

### Exercise 9

```{r NorthCarolina}
dn_nc <- dennys %>%
  filter(state == "NC")
nrow(dn_nc)

```
28 locations in North Carolina



```{r lqnc}
lq_nc <- laquinta %>%
  filter(state == "NC")
nrow(lq_nc)
```

12 La Quinta locations in NC


```{r joining2}
dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")
dn_lq_nc

```

336 observations, the variables are Address, city, zip, longitude, and latitude of both x and y (separately) and state



```{r distance2}
dn_lq_nc <- dn_lq_nc %>%
  mutate(dn_lq_nc, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_nc
```

```{r mindist2}
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_nc_mindist
```

```{r plot3}
dn_lq_nc%>%
  ggplot(mapping = aes(x=address.x, y=distance)) + geom_point()
```
```{r plot11}
dn_lq_nc_mindist%>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```

### Exercise 10


```{r Texas}
dn_tx <- dennys %>%
  filter(state == "TX")
nrow(dn_tx)

```
200 Dennys locations in Texas



```{r lqtx}
lq_tx <- laquinta %>%
  filter(state == "TX")
nrow(lq_tx)
```

237 La Quinta locations in NC


```{r joining3}
dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")
dn_lq_tx

```

47,400 observations, the variables are Address, city, zip, longitude, and latitude of both x and y (separately) and state



```{r distance3}
dn_lq_tx <- dn_lq_tx %>%
  mutate(dn_lq_tx, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_tx
```

```{r mindist3}
dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_tx_mindist
```

```{r plot10}
dn_lq_tx%>%
  ggplot(mapping = aes(x=address.x, y=distance)) + geom_point()
```
```{r plot4}
dn_lq_tx_mindist%>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```

### Exercise 11

```{r Pennsylvania}
dn_pa <- dennys %>%
  filter(state == "PA")
nrow(dn_pa)

```
40 locations in Pennsylvania



```{r lqpa}
lq_pa <- laquinta %>%
  filter(state == "PA")
nrow(lq_pa)
```

10 La Quinta locations in PA


```{r joining4}
dn_lq_pa <- full_join(dn_pa, lq_pa, by = "state")
dn_lq_pa

```

400 observations, the variables are Address, city, zip, longitude, and latitude of both x and y (separately) and state



```{r distance4}
dn_lq_pa <- dn_lq_pa %>%
  mutate(dn_lq_pa, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_pa
```

```{r mindist4}
dn_lq_pa_mindist <- dn_lq_pa %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_pa_mindist
```

```{r plot7}
dn_lq_pa%>%
  ggplot(mapping = aes(x=address.x, y=distance)) + geom_point()
```
```{r plot8}
dn_lq_pa_mindist%>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```

### Exercise 12
